Bayesian Hierarchical Model Predicts Dim Prospects for Trump's Victory in the 2024 U.S. Presidential Election

### 2024-11-02 paper/paper.qmd
llm usage updated for generation of reference reference updated for poll texas Appendix advantage and disadvantage of poll updated

Additionally, even with the best planning, The New York Times notes that they ensure the correct number of interviews by race, party, and region so that each NYT/Siena poll reaches, for example, the appropriate proportion of white Democrats from the West, or Hispanic Republicans from Maricopa County, Arizona [@nytimes2024methodology] [@nytimes2022wisconsin]. Still, some groups remain difficult to reach. Pollsters attempt to compensate for this by adjusting sample weights, but weighting itself introduces challenges. If certain groups are over- or under-adjusted within the sample, it may destabilize results and obscure potential sample limitations.

>>>>

Additionally, even with the best planning, the New York Times notes that they ensure the correct number of interviews by race, party, and region so that each NYT/Siena poll reaches, for example, the appropriate proportion of white Democrats from the West, or Hispanic Republicans from Maricopa County, Arizona [@nytimes2024methodology] [@nytimes2022wisconsin]. Still, some groups remain difficult to reach. Pollsters attempt to compensate for this by adjusting sample weights, but weighting itself introduces challenges. If certain groups are over- or under-adjusted within the sample, it may destabilize results and obscure potential sample limitations.

### The Advantages and Disadvantages of the Questionnaire Design

From polls like @nytimes2024texaspoll, the NYT/Siena polls maintain a high level of transparency, with each survey report publishing the questionnaire verbatim. This approach allows both the public and researchers to clearly understand the questions asked, enhancing trust. The questions are also straightforward, avoiding complex wording, which reduces the risk of misunderstandings across different respondent groups and improves response accuracy. Furthermore, the language is kept neutral, avoiding any leading phrases. This neutrality helps prevent bias that might arise from the phrasing of questions, ensuring that respondents’ answers are based on their genuine views rather than being influenced by question design, which supports the objectivity of the results.

The questionnaire is designed to examine respondents’ multifaceted characteristics, covering not only basic demographics like age, gender, and education level but also other personal traits and preferences. This information helps analyze political opinion differences across diverse backgrounds, making the results more representative. For example, the questionnaire may include questions on race, occupation, or top priorities when choosing a president—factors that often significantly impact voting preferences. It also includes questions about respondents' group characteristics, such as geographic location, community type (urban or rural), and political affiliation. By covering these factors, the survey can accurately reflect differences in public opinion across different regions and social environments. In a politically polarized climate, this detailed grouping helps identify regional or community-specific trends, showing how voters from different backgrounds feel about candidates or policies.

The questionnaire also includes questions about religious beliefs to analyze the impact of religion on voting intentions. Religious affiliation often aligns closely with political views, particularly in the U.S., where religious beliefs can strongly influence positions on morality, social policies, and party preferences. By incorporating religious affiliation, the questionnaire can better capture how religion influences voter choices, allowing for more targeted analysis of various religious backgrounds.

However, certain aspects of the questionnaire design have limitations. For instance, to keep questions concise, the survey may not delve deeply into complex topics, possibly overlooking subtle but important perspectives. For example, questions about voting likelihood are sometimes too general; each respondent may interpret “likelihood” differently based on their personal standards. Additionally, the lack of open-ended questions limits the depth of responses, reducing the questionnaire's ability to capture detailed opinions.

Another limitation is the focus on mainstream religions, often overlooking other groups such as Islam, Hinduism, Buddhism, and Sikhism. With increasing immigration, the number of voters from these religious backgrounds is rising, and their political preferences may differ from those of traditional religious groups. For instance, Muslim, Hindu, or Buddhist voters may prioritize issues such as cultural diversity, foreign policy, or religious freedom differently. The absence of questions covering these religions may result in a poll that does not fully reflect the priorities of all voters, limiting its representativeness. It also makes it harder to track the growing influence and political leanings of these groups. Without including these religious backgrounds, the questionnaire may miss important shifts in political attitudes among these emerging voter groups, potentially losing opportunities for future research.

The question in The Times/Siena Poll: Texas Topline, “If you had to decide between the two today, would you lean more toward” saw a high rate of “Don’t know/Refused” responses, at 66% [see @nytimes2024texaspoll]. This likely reflects a sense of dissatisfaction among voters due to political polarization. Many voters may feel discontent with both major candidates (Trump and Harris) and, as a result, choose not to respond when dissatisfied with the options or disillusioned with the political landscape. Additionally, the phrasing, “if you had to decide between the two today,” may make respondents feel pressured, leading to a refusal to answer. For those inclined toward third-party candidates or undecided voters, this type of question may further discourage responses. This question specifically targeted “leaners” (respondents who previously did not clearly support either Trump or Harris), often representing “swing” or “undecided” voters. Asking these voters, who are inherently unsure, to make a decision without strong support for either candidate could contribute to the high rate of refusals. A similar pattern is observed in questions like “(Leaners, if no candidate selected in initial multi-candidate question) If you had to decide today, would you lean more toward,” where respondents also frequently chose “Don’t know/Refused.”
